### 2020-09-18 latex/vlp.tex
Updated some phrasing on the VLP theorem
Updated intro to VLP section

%/========== Variable Length Pendulum ==========/%
\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
\section{Motivation}
The variable length pendulum (VLP) is a classical underactuated dynamical system
which is often used to model the motion of a person on a swing
\cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
The VLP also represents the swinging of a crane, the (simplified) motion of a
gymnast on a bar \cite{pendulum_length_giant_gymnastics}, and the
tuned-mass-damper systems which stabilize skyscrapers
\cite{vlp_tuned_mass_damper}.

The motion of the VLP has been well studied (see for instance
\cite{dynamics_periodic_vlp}), and many control mechanisms exist
to stabilize trajectories of the system. While many of these controllers 
are time-dependent, \cite{vlp_energy_shaping}
offers a time-independent technique to inject energy into the system and
stabilize desired energy level sets. The authors designed a controller through a
technique called \textit{energy shaping}, and proved that their control
mechanism would allow the VLP to achieve any desired energy level set.
However, the energy injection mechanism is
ad-hoc in the sense that it is not derived from natural behaviour. In
this chapter we will design VNHCs to inject energy into the VLP in a
time-independent way.
The difference compared to \cite{vlp_energy_shaping} is that our 
controller will maintain the structured motion of a human on a swing.

@@

\begin{thm}\label{thm:vlp-energy-stabilization}
   Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
   \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pm\pi,0)\right\}\).
   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
   \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
   \end{equation}
   with the property that the inequality is strict for almost every \(\theta\).
   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) leads to
   energy dissipation.
\end{thm}
\begin{proof}
    Choose, as a candidate anti-Lyapunov function, the energy for the average-length pendulum 
    \[
       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
                    + m g l_\text{avg} (1-\cos(q))
    \]
    which is positive definite at \((0,0)\) and has derivative 
    \[
      \dot{E}_\text{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
                 {l_\text{avg}^2l(\theta)^2}
    \]
    Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
    by Lemma \ref{lemma:sign-of-cube},
    \[ 
       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
    \]
    Then the derivative of \(E_\text{avg}\) is almost always positive, since
    \begin{align*}
       \label{eq:}
       \sign{\dot{E}_\text{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
                   &= -\sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                   &\geq 0 \text{ (by assumption)}
    \end{align*}
    Hence, \(E_\text{avg}\) is an anti-Lyapunov function.
    Since the initial condition is not an equilibrium,
    the orbit of the VNHC in the \((q,p)\) plane is diverging away from the
    origin and the VLP is gaining energy on average.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
    means \(E_\text{avg}\) is a regular Lyapunov function, since its derivative
    will be negative almost everywhere.
\end{proof}

>>>>

%/========== Variable Length Pendulum ==========/%
\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
\section{Motivation}
The variable length pendulum (VLP) is a classical underactuated dynamical system
which is often used to model the motion of a person on a swing
\cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
The VLP also represents the motion of the load at the end of a crane, 
the (simplified) motion of a gymnast on a bar
\cite{pendulum_length_giant_gymnastics}, and the tuned-mass-damper systems which
stabilize skyscrapers \cite{vlp_tuned_mass_damper}.

The motion of the VLP has been well studied (see for instance
\cite{dynamics_periodic_vlp}), and many control mechanisms exist
to stabilize trajectories of the system. While many of these controllers 
are time-dependent, \citet{vlp_energy_shaping}
offer a time-independent technique to inject energy into the system. 
They designed a controller through a technique called \textit{energy
shaping}, and proved that their control mechanism would allow the VLP to achieve
any desired energy level set.
However, the energy injection mechanism is ad-hoc in the sense that it is not
derived from natural behaviour. 
In this chapter we will design VNHCs to inject energy into the VLP in a
time-independent way.
The difference compared to \cite{vlp_energy_shaping} is that our 
controller will maintain the structured motion of a human on a swing.

@@

\begin{thm}\label{thm:vlp-energy-stabilization}
   Take the VLP with Hamiltonian dyanmics (\ref{eqn:vlp-hamiltonian}) and define
   \(\theta := \arctan_2(p,q)\). Suppose the initial condition is not one of the
   equilibria: \((q(0),p(0)) \not \in \left\{(0,0), (\pm\pi,0)\right\}\).
   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
   \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
   \end{equation}
   with the property that the inequality is strict for almost every \(\theta\).
   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) leads to
   energy dissipation.
\end{thm}
\begin{proof}
    Choose, as a candidate anti-Lyapunov function, the energy for the
    average-length pendulum (with zero potential at the bottom of the swing)
    \[
       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
                    + m g l_\text{avg} (1-\cos(q))
    \]
    which is positive definite at \((0,0)\) and has derivative 
    \[
      \dot{E}_\text{avg} = \frac{-g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
                 {l_\text{avg}^2l(\theta)^2}
    \]
    Observe that \(\sign{\sin(q)p} = \sign{\sin(2\theta)}\) and, 
    by Lemma \ref{lemma:sign-of-cube},
    \[ 
       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
    \]
    Then the derivative of \(E_\text{avg}\) is almost always positive, since
    \begin{align*}
       \label{eq:}
       \sign{\dot{E}_\text{avg}} &= \sign{-\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
                   &= -\sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                   &\geq 0 \text{ (by assumption)}
    \end{align*}
    Hence, \(E_\text{avg}\) is an anti-Lyapunov function.
    Since the initial condition is not an equilibrium,
    the orbit of the VNHC in the \((q,p)\) plane is diverging away from the
    origin and the VLP is gaining energy on average.
    Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
    means \(E_\text{avg}\) is a regular Lyapunov function, since its derivative
    will be negative almost everywhere.
\end{proof}
